---
title: "Investigating the Infraction Columns"
author: "Aaron Jacobs"
date: "January 22, 2017"
output:
  md_document:
    variant: markdown_github
    fig_width: 6
    fig_height: 4
    preserve_yaml: true
---

The [data dictionary](http://opendata.toronto.ca/revenue/parking/ticket/parking_tickets_readme.xls)
for the raw parking tickets data sets seems to indicate that there are three
columns containing the definition of the particular infraction:
`infraction_code`, `infraction_description`, and `set_fine_amount`. It sounds
like all three of these columns should be consistent across all tickets, so we
can significantly reduce the size of the data by moving the unique infraction
entries to their own table, and keeping only a reference to the infraction code
for each ticket entry.

However, we need to check that these columns really are that simple before
proceeding, so we load the data in the staging table.

```{r load-db}
library(dplyr, warn.conflicts = FALSE)

db <- src_sqlite("../tickets.sqlite")
tickets_staging <- tbl(db, "tickets_staging")
```

Now we can compute the frequency of all of unique combinations of these columns
in the raw data:

```{r infractions, cache = TRUE}
infractions <- tickets_staging %>%
    select(code = infraction_code, desc = infraction_description,
           fine = set_fine_amount) %>%
    group_by(code, desc, fine) %>%
    summarise(freq = n()) %>%
    ungroup() %>%
    collect()

infractions
```

OK, so there are two obvious problems with this from the outset. The first is
that it looks like at least some of the codes have multiple descriptions. The
second is that some of the codes seem to have different fines for different
entries. I'll return to this second problem below, but for now let's focus on
the first one.

The first thing to do is check how big a problem this is. What is the set of
codes that are "well-defined", in the sense that they have only one definition
in the dataset?

```{r well-defined}
well.defined <- infractions %>%
    # Pull out all of the codes that only have one "definition".
    group_by(code) %>%
    summarise(freq = n()) %>%
    ungroup() %>%
    filter(freq == 1) %>%
    select(code) %>%
    # Semi join on these well-defined infractions.
    semi_join(infractions, ., by = "code")

well.defined
```

Compare the number of rows in the well-defined group against the total number of
codes we've got:

```{r total-code-count}
infractions %>%
    summarise(total.codes = n_distinct(code))
```

So there are quite a few problematic definitions, although not too many to go
over completely manually if necessary.
