---
title: "Investigating the Location Columns"
author: "Aaron Jacobs"
date: "January 24, 2017"
output:
  md_document:
    variant: markdown_github
    fig_width: 6
    fig_height: 4
    preserve_yaml: true
---

The [data dictionary](http://opendata.toronto.ca/revenue/parking/ticket/parking_tickets_readme.xls)
for the raw parking tickets data sets points to a complex encoding of the
location the ticket was given at. It looks as though there can be two street
addresses, along with a description of the "proximity" of each. What does this
mean in practice? And is it possible to map this location encoding to, say,
longitude and latitude?

To get a sense of what is going on, we can load the data in the staging table

```{r load-db}
library(dplyr, warn.conflicts = FALSE)

db <- src_sqlite("../tickets.sqlite")
tickets_staging <- tbl(db, "tickets_staging")
```

And take a peek at the location rows:

```{r raw-locations}
tickets_staging %>%
    select(location1:location4) %>%
    print(n = 20)
```

From this particular sample, this data doesn't look so bad. The answer to why
there are sometimes one and sometimes two street addresses seems to be that some
entries point to an intersection.

How complete is the location data? And how often is an intersection -- as
opposed to a single address -- assigned to a ticket?

```{r location-col-breakdown, cache = TRUE}
location.counts <- tickets_staging %>%
    summarise(total = n(),
              one = sum(!is.na(location2)),
              two = sum(!is.na(location4))) %>%
    collect()

location.counts %>%
    mutate(one = one / total, two = two / total)
```

So these intersections are less common than one might have thought, looking at
the sample above. The overwhelmingly more common case is that of a single
address. It's also clear that the location data is pretty complete. More than 99
percent of tickets have non-missing data.
